# Insert chapter 2 title here
title: 'Chapter 2: Regression and Model Validation'
author: "Andrew House"
*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

First we read in the data and check it's structure and dimension. 

```{r}
learning2014 <- read.table('http://s3.amazonaws.com/assets.datacamp.com/production/course_2218/datasets/learning2014.txt', sep= ",", head =T )

dim(learning2014)
#166  7
str(learning2014)
### Data has 166 entires with 7 different variables. Data was collected in 2014-2015. 
date()
```


```{r}
library(ggplot2)

# initialize plot with data and aesthetic mapping
p1 <- ggplot(learning2014, aes(x = attitude, y = points, col = gender))

# define the visualization type (points)
p2 <- p1 + geom_point()

# draw the plot


# add a regression line
p3 <- p2 + geom_smooth(method = "lm")

# add a main title and draw the plot
p4 <- p3 + ggtitle("Student's attitude versus exam points")
p4
```
```{r}
### plot the variable and look for correlation between the variable
library(GGally)
library(ggplot2)

# create a more advanced plot matrix with ggpairs()
p <- ggpairs(learning2014, mapping = aes(col = gender), lower = list(combo = wrap("facethist", bins = 20)))
p
```



Here we go again...
