# Chapter 3
title: 'Chapter 3: Logistic Regression'
author: "Andrew House"
*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

First we read in the data and check it's structure and dimension.

```{r}
#Reading the data
pormath <- read.table('https://github.com/rsund/IODS-project/raw/master/data/alc.csv',sep= ",", head =T )


#Structure and dimensions of the data
str(pormath)
```


Variables not used for joining the two data have been combined by averaging (including the grade variables)
'alc_use' is the average of 'Dalc' and 'Walc'
'high_use' is TRUE if 'alc_use' is higher than 2 and FALSE otherwise

```{r}
library(tidyr); library(dplyr); library(ggplot2)

# glimpse at the alc data
#glimpse(alc) 

# use gather() to gather columns into key-value pairs and then glimpse() at the resulting data
#gather(alc) %>% glimpse

# draw a bar plot of each variable
#gather(alc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()

```
My personal hypotheses

Guardian
freetime 
go out
internet 

```{r}
#alc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_grade = mean(G3))
pormath %>% group_by(sex, high_use) %>% summarise(count = n(), mean_studytime = mean(studytime))
```


```{r}
pormath %>% group_by(sex, high_use) %>% summarise(count = n(), mean_absences = mean(absences))
```


```{r}
pormath %>% group_by(sex, high_use) %>% summarise(count = n(), mean_freetime = mean(freetime))
```


```{r}
pormath %>% group_by(sex, high_use) %>% summarise(count = n(), mean_goout = mean(goout))
```


```{r}
library(cowplot)
#Initializing a barplot of age
g1 <- ggplot(pormath, aes(x = absences, fill = sex)) + geom_bar() #+ ylab("age")

#... of going out with friends (goout)
g2 <- ggplot(pormath, aes(x = goout, fill = sex)) + geom_bar()

#... of absences
g3 <- ggplot(pormath, aes(x = studytime, fill = sex)) + geom_bar()

#... of final grade (G3)
g4 <- ggplot(pormath, aes(x = freetime, fill = sex)) + geom_bar()

#... of alcohol consumption (alc_use)
g5 <- ggplot(pormath, aes(x = alc_use, fill = sex)) + geom_bar()

#... of whether the consumption is high/low (high_use)
g6 <- ggplot(pormath, aes(x = high_use, fill = sex)) + geom_bar()


#Plotting all 4 barplots
plot_grid(g1, g2, g3, g4, g5, g6, labels="AUTO")
```

