# Insert chapter 3 title here
title: 'Chapter 3: Logistic Regression'
author: "Andrew House"
*Describe the work you have done this week and summarize your learning.*

- Describe your work and results clearly. 
- Assume the reader has an introductory course level understanding of writing and reading R code as well as statistical methods.
- Assume the reader has no previous knowledge of your data or the more advanced methods you are using.

First we read in the data and check it's structure and dimension.

```{r}
#Reading the data
pormath <- read.table('https://github.com/rsund/IODS-project/raw/master/data/alc.csv',sep= ",", head =T )


#Structure and dimensions of the data
str(pormath)
```





```{r}
library(tidyr); library(dplyr); library(ggplot2)

# glimpse at the alc data
#glimpse(alc) 

# use gather() to gather columns into key-value pairs and then glimpse() at the resulting data
#gather(alc) %>% glimpse

# draw a bar plot of each variable
#gather(alc) %>% ggplot(aes(value)) + facet_wrap("key", scales = "free") + geom_bar()

```


```{r}
#alc %>% group_by(sex, high_use) %>% summarise(count = n(), mean_grade = mean(G3))

```

